Statistical Modeling of Galaxy Catalogues with Normalizing Flows

### 2022-11-30 src/tex/ms.tex
Updated appendix on training.

We train an ensemble of 4 normalizing flows, each with the same architecture and training schedule as the flow described in Section \ref{sec:galaxy-catalog}.
PZFlow makes this easy to do automatically, by simply swapping \texttt{FlowEnsemble} for \texttt{Flow}.

@@

\begin{figure*}[t!]
    \script{forward_model/plot_galaxy_losses.py}
    \begin{centering}
        \includegraphics{figures/galaxy_flow_losses.pdf}
        \caption{
            Training losses for the galaxy flows.
            After epoch 50 and 100, you can see a drop in the loss due to the decrease in the learning rate.
        }
        \label{fig:galaxy-flow-losses}
    \end{centering}
\end{figure*}

\note{the main flow:}
We trained the normalizing flow for 150 epochs using the Adam optimizer \citep{adam} (with a learning rate of $10^{-3}$, that decreased by a factor of 10 every 50 epochs) to minimize the negative log-likelihood,
\begin{align}
    \mathcal{L} = - \, \mathbb{E}[ \, \log p(x) \, ],
\end{align}
where the expectation is performed over galaxies in the training set and $p(x)$ is defined in Equation \ref{eq:px}.
Training took 7 minutes on a Tesla P100 12GB GPU.

\note{
    Need to add the curves for the ensembles.
    Move discussion of epochs and learning rates here too?    
}

>>>>

We train an ensemble of 4 normalizing flows, each with the same architecture and training schedule as the flow described in Section \ref{sec:galaxy-catalog}.
PZFlow makes this easy to do automatically, by simply swapping \texttt{FlowEnsemble} for \texttt{Flow} in the code.

@@

\begin{figure*}[t!]
    \script{forward_model/plot_galaxy_losses.py}
    \begin{centering}
        \includegraphics{figures/galaxy_flow_losses.pdf}
        \caption{
            Training losses for the galaxy flows.
            Left: losses for the normal flow.
            After epochs 50 and 100, you can see a drop in the loss due to the decrease in the learning rate.
            Right: losses for the conditional flow.
            After epochs 150 and 300, you can see a drop in the loss due to the decrease in the learning rate.
        }
        \label{fig:galaxy-flow-losses}
    \end{centering}
\end{figure*}

\begin{figure}[t!]
    \script{photo-z/plot_ensemble_losses.py}
    \begin{centering}
        \includegraphics{figures/ensemble_losses.pdf}
        \caption{
            Training losses for the four flows in the flow ensemble.
            You can see that each of the flow achieves a different minimum loss.
            After epochs 50 and 100, you can see a drop in the loss due to the decrease in the learning rate.
        }
        \label{fig:ensemble-losses}
    \end{centering}
\end{figure}

In this section we list some technical details of training the normalizing flows.
Every flow is trained via minimizing the negative log-likelihood
\begin{align}
    \mathcal{L} = - \, \mathbb{E}[ \, \log p(x) \, ],
\end{align}
where the expectation is performed over galaxies in the training set and $p(x)$ is defined in Equation \ref{eq:px}.

For the main flow in Section \ref{sec:galaxy-catalog}, we trained for 150 epochs.
We used the Adam optimizer \citep{adam}, starting with a learning rate of $10^{-3}$.
We decreased the learning rate by a factor of 10 every 50 epochs.
Training took 7 minutes on a Tesla P100 12GB GPU.
The training loss for this flow is in the left panel of Figure \ref{fig:galaxy-flow-losses}.

For the conditional flow in \ref{sec:galaxy-catalog}, we trained for 450 epochs.
Again, we used Adam with an initial learning rate of $10^{-3}$.
We decreased the learning rate by a factor of 10 every 150 epochs.
The training loss for this flow is in the right panel of Figure \ref{fig:galaxy-flow-losses}.

For each of the flows that make up the flow ensemble in \ref{sec:photo-z}, we trained for 150 epochs using the Adam optimizer.
We started each with a learning rate of $10^{-4}$, and decreased the learning rate by a factor of 10 every 50 epochs.
The training loss for the ensemble is in Figure \ref{fig:ensemble-losses}.
You can see that each of the flows achieves a different minimum loss.
Apparently, each has found a different potential solution in the neural network's parameter space.
